Best Practices for Alchemical Free Energy Calculations: v\versionnumber

### 2019-12-19 paper/manuscript.tex
edits on contribs

\newcommand{\versionnumber}{0.1}  % you should update the minor version number in preprints and major version number of submissions.
\newcommand{\githubrepository}{\url{https://github.com/michellab/alchemical-best-practices}}  %this should be the main github repository for this article
\newcommand{\expect}[1]{\left\langle{#1}\right\rangle}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% ARTICLE SETUP
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\title{Best Practices for Alchemical Free Energy Calculations: v\versionnumber}
\author[2*]{Antonia S. J. S. Mey}
\author[7]{Bryce K. Allen}
\author[1*]{John D. Chodera}
\author[2,12]{Maximilian Kuhn}
\author[2*]{Julien Michel}
\author[3*]{David L. Mobley}
\author[1]{Levi N. Naden}
\author[4]{Samarjeet Prasad}
\author[5]{Julia E. Rice}
\author[1,9]{Andrea Rizzi}
\author[2]{Jenke Scheen}
\author[6*]{Michael R. Shirts}
\author[11]{Gary Tresadern}
\author[10]{Huafeng Xu}

@@

\begin{figure}
    \includegraphics[width=0.95\linewidth]{figures/fig1_what_is_alchemy/Figure.pdf}   
    \caption{\textbf{Illustration of the types of free energies differences that can be calculated using alchemical free energy methods.} A: Change in free energy due to a conformational changes of the moelcule. B: Partition coefficient such as $\log P$ or $\log D$ can be computed using alchemical free energy methods. Here, as an example the partition coefficient between methanol and water is shown. C: Free energy difference associated with the insertion of a molecule into a membrane. D: Effect of mutations of protein or host residues on free energies of binding.  E: Absolute free energies of binding of a small molecule to a host (e.g. protein), F: Relative free energy of binding of one molecule with respect to another, here toluene and benzyl alcohol.
    \label{fig:fig_what_is_alchemy}
    }
\end{figure}

%%%%%%%%%%%%%%%%%%%%
% Prerequesites    %
%%%%%%%%%%%%%%%%%%%%
\section{Prerequisites and Scope}
\label{sec:pre}
%Here you would identify prerequisites/background knowledge that are assumed by your work and your checklist which you view as critical, ideally giving links to good sources on these topics.
%Checklists are normally focused on errors made by users with training and experience in molecular simulations, so you can assume a basic familiarity with the fundamentals of molecular simulations.
\todo[inline, color=green!20]{ASJSM: @everyone This section is up for review to also check flow with intro and rest of document.}
This guide is meant to serve as a comprehensive starting point for new practitioners, as well as a reference guide with a convenient checklist (Section~\ref{sec:checklist}) for more seasoned practitioners, to help with standard alchemical simulation and analysis practices. It can also help as a set of best practices references for reviewing publications. 
This guide aims to answer the following questions:
\begin{itemize}
    \item Is my problem suitable for an alchemical free energy calculation? 
    \item How do I choose the right free energy protocol and run it? How do I analyze my simulations accurately? 
    \item What software tools are available to simulate alchemical free energy protocols? 
\end{itemize}

@@

\todo[inline,color={green!20}]{ASJSM: Discuss implications of the SAMPL6 SAMPLing paper~\cite{rizzi2019sampl6} on reproducibility between simulation packages and potential sensitivity to other calculation details. @jmichel, @dmobley}
Finite computing resources necessarily limit the generated number of uncorrelated samples of potential energy surfaces, and therefore alchemical free energy calculations only give free energy estimates to within a given precision. An important consideration is how reproducible alchemical free energy calculations are in practice. In simple cases such as absolute hydration free energies of small organic molecules, or relative hydration free energy calculations between structurally similar organic molecules, it should be possible to obtain highly precise estimates with a given software package (with a sample standard deviation under 0.01 kcal/mol). For more complex use cases such as protein-ligand binding free energies the repeatability is often substantially worse. A good practice is to perform two or three runs of the same perturbation to assess repeatability with a given protocol. The sample standard deviation will give a crude estimate of the reliability of the estimates, and whether the precision is sufficient for the problem at hand. Where this is practical, a more stringent test is to use different input coordinates for each repeat run. Note that these issues concern calculations carried out with a single software package. The issue of reproducibility of free energy calculations across different simulation packages has attracted attention recently. Greater variability is expected due to methodological differences such as integrators, thermostats, barostats, treatment of long-range electrostatics. For absolute and relative hydration free energies of small organic molecules a variability of ca. 0.2 kcal/mol between popular simulation packages has been reported.(REF Reproducibility of hydration free energies Loefler et al.  JCTC 2018) In the recent SAMPL6 SAMPLing challenge a larger variability of 0.3 to 1.0 kcal/mol was noted in the computed absolute binding free energies of host/guest systems even though similar input and simulation parameters were used (Ref Rizzi et al SAMPLing biorxiv). Further work is needed to ensure reproducibility of alchemical free energy calculations across implementations to guarantee that force-field development efforts lead to transferable potential energy functions.

@@

Here we include a simple estimate of the value provided by alchemical
free energy calculations in lead optimization.  Let $P(\Delta\Delta
G)$ be the probability distribution of the changes in the binding free
energies of a new set of molecules during one round of lead
optimization, and let $P(\Delta\Delta G^\dagger|\Delta\Delta G)$ be the
conditional probability of the binding free energy change computed by
the free energy calculations, $\Delta\Delta G^\dagger$, given the actual
change $\Delta\Delta G$.  The latter conditional probability can be modeled
by a normal distribution
\begin{equation}
  P(\Delta\Delta G^\dagger|\Delta\Delta G) = \frac{1}{\sqrt{2\pi\sigma^2}}
  \exp\left(-\frac{(\Delta\Delta G^\dagger - \Delta\Delta G)^2}{2\sigma^2}\right),
  \label{eqn:free-energy-distribution}
\end{equation}
where $\sigma$ signifies the accuracy of free energy calculations.
Here we assume that there is no systematic bias in the free energy
calculations, i.e., on average, the free energy change computed by
free energy calculations agrees with the actual free energy change.
In lead optimization guided by free energy calculations, we only
synthesize and experimentally test molecules that are predicted to
have favorable free energy changes.  We are thus interested in how
often that a molecule predicted to bind stronger actually turns out to
bind stronger.  In other words, we are interested in the conditional
probability:
\begin{equation}
  P(\Delta\Delta G<0|\Delta\Delta G^\dagger<0).
  \label{eqn:true-positive}
\end{equation}
For illustrative purposes, we assume that the actual changes in the
binding free energies for a set of new molecules also follow normal
distribution, that the standard deviation in the changes is $RT\ln 5$
(corresponding to a 5-fold change in the binding affinities), and that
1 in 10 new molecules have increased binding affinity ($\Delta\Delta G
\leq 0$).  Under such assumptions, the conditional probability in
Eq.~\ref{eqn:true-positive} can be easily computed.  If the accuracy
of free energy calculations is $\sigma = 1$ kcal/mol, $P(\Delta\Delta
G<0|\Delta\Delta G^\dagger<0) = 0.35$, which means that out of every
10 molecules selected for predicted favorable free energy change, on
average 3.5 molecules will have actual favorable free energy change.
In other words, selection by free energy calculations yields 3.5 times
more molecules of improved affinities than selection without free
energy calculations.
Available computational resources and timescales of motion also factor
into this initial analysis. An individual free energy calculation
involves simulations at many different intermediate states (perhaps
20-40 or more) and each of these must typically be long enough to
capture the relevant motions in the system. If such motions are
microsecond events or longer, the computational cost of running 20-40
microsecond or longer simulations for each of $N$ ligands may become
prohibitive, or at least should be carefully considered. 
\subsection*{Can I afford the calculation?}
Furthermore, are available computational resources sufficient that throughput will be
reasonable compared to needs of experimental collaborators working on
this system? How many ligands ($N$) can you afford to handle given
your computational resources?
This analysis should be done up front, part of ``counting the cost''
of involvement in a particular project. In some cases, the analysis may conclude that free energy calculations will not be feasible for the proposed problem.
\todo[inline, color={green!20}]{ASJSM: @everyone: Might be nice to do an example calculation X perturbation on GTX 1080s on a AWS cluster cost \$\$X. Synthesis costs \$\$Y. }
\subsection*{Is an exploratory study what I want?}
An additional consideration is how much is known about your particular
target, ligand binding modes in the target, and any relevant motions
-- essentially, has it been studied enough to know whether it might be
suitable for free energy calculations? It is important to know if the system has hardly been studied, because should the initial calculations perform poorly, the effort may turn into an attempt to understand the relevant sampling, force field, or system preparation problems.
If you are unsure whether your project is feasible, one option may be
to conduct a short exploratory study to assess tractability for just a very small
number of ligands. This can be sufficient to get an initial
idea of feasibility and accuracy of the calculations for the
proposed target.

@@

It is expected that AFE methods can provide a valuable tool for drug discovery, capable of outperforming alternative approaches. (CITE Kuhn et al J. Med. Chem. 2017, 60, 6, 2485-2497). It has been stressed that successful application in industry requires more knowledge of the "domain of applicability", in other words, a greater understanding of the success and failure~\cite{sherborne2016collaborating}, which in turn requires robust protocols for preparing, submitting and analysing AFE calculations. In this regard, the issues mentioned in the previous section such as understanding the suitability and timescales to capture the structure activity relationships (SAR), and performing up-front tests of performance are all relevant to drug discovery applications. Without venturing too far into details of system setup, which is beyond the scope of this article, there are some additional points to consider. 

@@

The calculations aim to capture the alchemical change from one ligand to another as accurately as possible. Therefore, it is necessary to consider details of the experimental setup, such as pH. Biological assays are usually run at neutral pH but this is not always the case, and protein and ligand preparation protocols should reflect this. In more detail, the formal charge, or tautomeric state of the small molecules can change within a series of analogs. The change of pKa is a common medicinal chemistry strategy to modify drug properties and would normally require explicit effort to incorporate into alchemical calculations. Further concerns about the experimental setup require understanding what protein construct is used in the bioassay. For instance, does the X-ray that is to be used for the calculations match the construct used for screening (i.e. catalytic vs full length, monomer vs dimer) (cite Perez-Benito Sci Rep 2018, 8, 4883)? Also, were certain co-factors or partner proteins required in the bioassay? 

@@

Running binding FE calculations in a drug discovery application will typically require the use of software or tools to facilitate the calculations of many compounds. The commercial implementation from Schr\"{o}dinger, FEP+, uses a graphical user interface and once the user has calculated any missing force field parameters, calculations can be prepared and submitted within minutes. The software builds the network of perturbations, connecting similar compounds based on LOMAP methodology~\cite{liu2013lead}, all inputs for the Desmond simulations are built automatically and the job can be saved or run immediately on available GPUs. The cost of this software is often prohibitive but alternative tools are becoming available to allow relatively straight forward running of alchemical FE calculations with other software~\cite{gapsys2015pmx, loeffler2015fesetup} (CITE New Merz paper, cite PMX paper Vytas and Bert de Groot, cite FESETUP, cite FEanalysis). One consideration at this point is to define the perturbations suitable for the application. For instance, if the goal is to accurately assess the relative binding energy of a small number of compounds, possibly with challenging synthesis, the map of perturbations may contain as many connections between compounds as affordable. However, when running calculations on hundreds of compounds a so called ‘star-map’ can be used that just contains one connection per compound: perturbing every compound to a central ligand, typically the crystal structure ligand. In this way the top-ranking examples can be readily identified and submitted to additional calculations in a second round. 
%
\subsection*{Making predictions, understanding errors}
For prospective drug discovery applications there are several other considerations including understanding likely errors and taking selection bias into account. It is crucial when proposing compounds for synthesis to have some idea of the underlying error or uncertainty in the predictions. Fortunately, there are several useful indicators, results suggest MBAR errors for perturbations greater than 0.3 kcal/mol can be indicative of inaccuracy~\cite{perez-benito2019predicting}. In addition, the hysteresis can be checked by comparing the difference for forward and reverse perturbations of the same structural change, or in the case of large maps of perturbations cycle closure hysteresis~\cite{wang2013modeling} indicates problematic perturbations involved in cycles connecting many compounds. Selection bias, as described above, can be problematic in prospective applications as there is normally an aim to focus on only a narrow range of predicted activity (i.e. the most active). Thus, correlation between prediction and subsequent experiment may lead to disappointment but corrections can be applied based on previous recommendations~\cite{abel2017critical}. 
%

In summary, the successful use of alchemical calculations in drug discovery requires working in the domain of applicability, using a high quality X-ray structure of the target compounds, and testing the approach retrospectively to ensure the system setup is well-behaved. Always assess the confidence in the predictions and communicate this when discussing with experimentalists. Consider performing repeat calculations for at least some of the perturbations in the study. In a drug discovery setting it is inevitable to consider larger virtual libraries, in these cases again evaluate if shorter simulation time per $\lambda$ window can be sufficient for a ‘virtual screening’ mode of application. It is expected that future tools may allow efficient exploration of large virtual libraries. There are many accounts of success of alchemical FE calculations, the methods show good performance towards the goal of binding energy prediction. However, it is important to have realistic expectations. Structure based drug design projects are often capable of improving potency relatively quickly and the range of activity narrows to just two-to-three log units. It may seem hard to impact with substantially different, more potent, stand-out compounds in this scenario, but binding energy prediction can still be extremely useful for ensuring activity is maintained as other properties are optimized. An interesting cost benefit analysis has been performed, see drug discovery strategy related discussions of Mobley Abel and co. From a drug discovery point of view, FEP is expanding its domain of applicability, there are reports of success using homology models and GPCRs for instance, as well as enabling charge change and scaffold hopping, but these systems are undoubtedly more difficult.  In the meantime the use cases are expanding to resistance prediction, disease mutation prediction, solubility prediction – an exciting future for alchemical FE calculations. 
%
%
%
%%%%%%%%%%%%
% Step 1   %
%%%%%%%%%%%%
%
\section{Simulation prerequisites}
\label{sec:step1}

The alchemical free energy protocols, that are discussed in Section~\ref{sec:step2}, are defined for a specific type of free energy calculation, i.e. a free energy of binding or a free energy of hydration. Depending on the type of simulations different considerations have to be made for ligands, solvent, and host molecules (in the case of the estimation of free energies of binding).
\subsection*{Free energies of binding}
In principle in the limit of sufficient conformational sampling, the free energy changes estimated from an alchemical free energy calculation should be independent of the system's initial coordinates. However, in practice because simulations of thermodynamic states are of finite duration (typically currently 1-100 ns per state) this is only true for certain classes of alchemical free energy calculations such as relative or absolute free energies of hydration of small and relatively rigid organic molecules. Host-guest or protein-ligand complexes typically exhibit slowly relaxing degrees of freedom that exceed significantly the duration of an alchemical free energy calculation. It is therefore generally important to carefully select input coordinates to obtain satisfactory results. 
The following questions may be relevant before diving into the simulation setup. 
\begin{itemize}
    \item Do I have one or multiple good host structures? (e.g. a good resolution X-ray crystal of the protein target)
    \item Should I include buried waters, or other small molecules that can be found in an X-ray structure.
    \item Are my ligands part of a congeneric series? (simple R group substitutions around the same scaffold)
    \item Do I have information on one or all of the ligands binding sites (e.g. a X-ray structure)
\end{itemize}

@@

Many relative free energy planning tools have been set up to avoid changing the net charge of the systems considered, including LOMAP~\cite{} and Schr\"{o}dinger's FEP+~\cite{chen2018accurate}. Absolute free energy calculations can also potentially avoid changing the charge of the system by making a charge perturbation of equal and opposite sign elsewhere in the system; for example, as a charged ligand is removed, a charged counterion of opposite sign could also be removed, or one of the same sign could be inserted. This approach is sometimes referred to as an "alchemical ion" approach to dealing with the needed charge change, and is also employed by the Yank free energy package~\cite{wang2013identifying}.

@@

Before launching alchemical free energy calculations it is wise to consider how convergence and completion will be assesses. Different conditions on when to stop alchemical free energy calculations should be determined, and this may require several iterative checks and therefore modifications to the calculation protocol.
One useful metric to use for termination is the expected or desired uncertainty of a desired free energy estimate, though care must be exercised should the uncertainty estimate prove unreliable.
In particular, if the rate of change in the free energy estimate is significant when this condition is met, the simulation may not be locally converged, and more sampling may be necessary to determine a stable free energy estimate which is no longer changing significantly over time. 
However, this is not the only metric which should be used, as the uncertainty only captures the information about the sampled phase space, not necessarily the entirety of the phase space.  
For example, 
convergence of relative free energy calculations in predictive simulations where the entire phase space is not known in advance, requires sampling the different kinetically stable states \todo[inline]{cite Mobley and Klimovich JCP 2012}. 
This highlights the importance of choosing the correct thermodynamic path to ensure you sample the required thermodynamic states as discussed in Section~\ref{sec:important_path}.

@@

\paragraph{Subsampling data to generate uncorrelated samples}
Once the statistical inefficiency $g$ has been estimated, it is straightforward to subsample the correlated timeseries simulation data to produce effectively uncorrelated data that can ba fed to estimators of the free energy.
Suppose the correlated timeseries is $\{a_t\}_{t=1}^T$; we can form a new timeseries of $N_{\mathrm{eff}} \approx T / g$ effectively uncorrelated samples by selecting a subset of indices $\{ \: t = \mathrm{round}((n-1) \, g) \: | \: n \in \mathrm{range}(1,\ldots ,N) \: \}$ where $\mathrm{round(x)}$ denotes rounding to the nearest integer.

>>>>

\newcommand{\versionnumber}{0.1}  % you should update the minor version number in preprints and major version number of submissions.
\newcommand{\githubrepository}{\url{https://github.com/michellab/alchemical-best-practices}}  %this should be the main github repository for this article
\newcommand{\expect}[1]{\left\langle{#1}\right\rangle}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% ARTICLE SETUP
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\title{Best Practices for Alchemical Free Energy Calculations: v\versionnumber}
\author[2*]{Antonia S. J. S. Mey}
\author[7]{Bryce K. Allen}
\author[1*]{John D. Chodera}
\author[2,12]{Maximilian Kuhn}
\author[2*]{Julien Michel}
\author[3*]{David L. Mobley}
\author[1]{Levi N. Naden}
\author[4]{Samarjeet Prasad}
\author[5]{Julia E. Rice}
\author[1,8]{Andrea Rizzi}
\author[2]{Jenke Scheen}
\author[6*]{Michael R. Shirts}
\author[11]{Gary Tresadern}
\author[10]{Huafeng Xu}

@@

\begin{figure}
    \includegraphics[width=0.95\linewidth]{figures/fig1_what_is_alchemy/Figure.pdf}   
    \caption{\textbf{Illustration of the types of free energies differences that can be calculated using alchemical free energy methods.} A: Change in free energy due to a conformational change of the molecule. B: Partition coefficient such as $\log P$ or $\log D$ can be computed using alchemical free energy methods. Here, as an example the partition coefficient between methanol and water is shown. C: Free energy difference associated with the insertion of a molecule into a membrane. D: Effect of mutations of protein or host residues on free energies of binding.  E: Absolute free energies of binding of a small molecule to a host (e.g. protein), F: Relative free energy of binding of one molecule with respect to another, here toluene and benzyl alcohol.
    \label{fig:fig_what_is_alchemy}
    }
\end{figure}

%%%%%%%%%%%%%%%%%%%%
% Prerequesites    %
%%%%%%%%%%%%%%%%%%%%
\section{Prerequisites and Scope}
\label{sec:pre}
%Here you would identify prerequisites/background knowledge that are assumed by your work and your checklist which you view as critical, ideally giving links to good sources on these topics.
%Checklists are normally focused on errors made by users with training and experience in molecular simulations, so you can assume a basic familiarity with the fundamentals of molecular simulations.
\todo[inline, color=green!20]{ASJSM: @everyone This section is up for review to also check flow with intro and rest of document.}
This guide is meant to serve as a comprehensive starting point for new practitioners, as well as a reference guide with a convenient checklist (Section~\ref{sec:checklist}) for more seasoned practitioners, to help with standard alchemical simulation and analysis practices. It can also help as a set of best practices which reviewers may wish to consider as they evaluate papers.

This guide aims to answer the following questions:
\begin{itemize}
    \item Is my problem suitable for an alchemical free energy calculation? 
    \item How do I choose the right free energy protocol and run it? \item How do I analyze my simulations accurately? 
    \item What software tools are available to simulate alchemical free energy protocols? 
\end{itemize}

@@

Finite computing resources necessarily limit the generated number of uncorrelated samples of potential energy surfaces, and therefore alchemical free energy calculations only give free energy estimates to within finite precision. An important consideration is how reproducible alchemical free energy calculations are in practice. In simple cases such as absolute hydration free energies of small organic molecules, or relative hydration free energy calculations between structurally similar small organic molecules, it should be possible to obtain highly precise estimates with a given software package (with a sample standard deviation under 0.01 kcal/mol). For more complex use cases such as protein-ligand binding free energies the repeatability is often substantially worse. A good practice is to perform two or three runs of the same perturbation to assess repeatability with a given protocol. The sample standard deviation will give a crude estimate of the reliability of the estimates, and whether the precision is sufficient for the problem at hand. Where this is practical, a more stringent test is to use different input coordinates for each repeat run.

Note that these issues concern calculations carried out with a single software package, but simulation package variations can introduce additional issues. Such issues of reproducibility of free energy calculations across different simulation packages have attracted attention recently. Greater variability is expected due to methodological differences such as integrators, thermostats, barostats, treatment of long-range electrostatics, and potentially other factors. For absolute and relative hydration free energies of small organic molecules a variability of ca. 0.2 kcal/mol between popular simulation packages has been reported. \cite{REF Reproducibility of hydration free energies Loefler et al.  JCTC 2018} In the recent SAMPL6 SAMPLing challenge a larger variability of 0.3 to 1.0 kcal/mol was noted in the computed absolute binding free energies of host/guest systems even though similar input and simulation parameters were used \cite{Ref Rizzi et al SAMPLing biorxiv} and, in many cases, single-point energies were identical or nearly so. Further work is needed to ensure reproducibility of alchemical free energy calculations across implementations to guarantee that force-field development efforts lead to transferable potential energy functions.

@@

Here we include a simple estimate of the value provided by alchemical
free energy calculations in lead optimization.  Let $P(\Delta\Delta
G)$ be the probability distribution of the changes in the binding free
energies of a new set of molecules during one round of lead
optimization, and let $P(\Delta\Delta G^\dagger|\Delta\Delta G)$ be the
conditional probability of the binding free energy change computed by
the free energy calculations, $\Delta\Delta G^\dagger$, given the actual
change $\Delta\Delta G$.  The latter conditional probability can be modeled
by a normal distribution
\begin{equation}
  P(\Delta\Delta G^\dagger|\Delta\Delta G) = \frac{1}{\sqrt{2\pi\sigma^2}}
  \exp\left(-\frac{(\Delta\Delta G^\dagger - \Delta\Delta G)^2}{2\sigma^2}\right),
  \label{eqn:free-energy-distribution}
\end{equation}
where $\sigma$ signifies the accuracy of free energy calculations.
Here we assume that there is no systematic bias in the free energy
calculations, i.e., on average, the free energy change computed by
free energy calculations agrees with the actual free energy change.
%
In lead optimization guided by free energy calculations, we only
synthesize and experimentally test molecules that are predicted to
have favorable free energy changes.  We are thus interested in how
often that a molecule predicted to bind stronger actually turns out to
bind stronger.  In other words, we are interested in the conditional
probability:
\begin{equation}
  P(\Delta\Delta G<0|\Delta\Delta G^\dagger<0).
  \label{eqn:true-positive}
\end{equation}
%
For illustrative purposes, we assume that the actual changes in the
binding free energies for a set of new molecules also follow normal
distribution, that the standard deviation in the changes is $RT\ln 5$
(corresponding to a 5-fold change in the binding affinities), and that
1 in 10 new molecules have increased binding affinity ($\Delta\Delta G
\leq 0$).  Under such assumptions, the conditional probability in
Eq.~\ref{eqn:true-positive} can be easily computed.  If the accuracy
of free energy calculations is $\sigma = 1$ kcal/mol, $P(\Delta\Delta
G<0|\Delta\Delta G^\dagger<0) = 0.35$, which means that out of every
10 molecules selected for predicted favorable free energy change, on
average 3.5 molecules will have actual favorable free energy change.
In other words, selection by free energy calculations yields 3.5 times
more molecules of improved affinities than selection without free
energy calculations.
%  
Available computational resources and timescales of motion also factor
into this initial analysis. An individual free energy calculation
involves simulations at many different intermediate states (perhaps
20-40 or more) and each of these must typically be long enough to
capture the relevant motions in the system. If such motions are
microsecond events or longer, the computational cost of running 20-40
microsecond or longer simulations for each of $N$ ligands may become
prohibitive, or at least should be carefully considered. 
%
\subsection*{Can I afford the calculation?}
Furthermore, are available computational resources sufficient that throughput will be
reasonable compared to needs of experimental collaborators working on
this system? How many ligands ($N$) can you afford to handle given
your computational resources? As cloud computing becomes more available, in-house GPU clusters may not be necessary if calculations are not run on a regular basis.
This analysis should be done up front, part of ``counting the cost''
of involvement in a particular project. In some cases, the analysis may conclude that free energy calculations will not be feasible for the proposed problem.
\todo[inline, color={green!20}]{ASJSM: @everyone: Might be nice to do an example calculation X perturbation on GTX 1080s on a AWS cluster cost \$\$X. Synthesis costs \$\$Y. }

Here, by ``cost'', we refer not just to financial cost of the calculations relative to experiments, but also time -- can the calculations can be run faster than experiments are done? How will the relevant resource and opportunity costs factor in? Both computation and experiment require human time, supplies (of different sorts), and equipment. In the extreme limit, for example, it would not make sense to spend a month running a binding free energy calculation if the equivalent experiment could be done in a day with resources already on hand. Such issues should be considered before deciding to conduct binding free energy calculations.
%
\subsection*{Is an exploratory study what I want?}
An additional consideration is how much is known about your particular
target, ligand binding modes in the target, and any relevant motions
-- essentially, has it been studied enough to know whether it might be
suitable for free energy calculations? It is important to know if the system has hardly been studied, because should the initial calculations perform poorly, the effort may turn into an attempt to understand the relevant sampling, force field, or system preparation problems.
%
If you are unsure whether your project is feasible, one option may be
to conduct a short exploratory study to assess tractability for just a very small
number of ligands. This can be sufficient to get an initial
idea of feasibility and accuracy of the calculations for the
proposed target.

@@

It is expected that AFE methods can provide a valuable tool for drug discovery, capable of outperforming alternative approaches. (CITE https://pubs.acs.org/doi/abs/10.1021/acs.jmedchem.6b01881). It has been stressed that successful application in industry requires more knowledge of the "domain of applicability", in other words, a greater understanding of the success and failure~\cite{sherborne2016collaborating}, which in turn requires robust protocols for preparing, submitting and analysing AFE calculations. In this regard, the issues mentioned in the previous section such as understanding the suitability and timescales to capture the structure activity relationships (SAR), and performing up-front tests of performance are all relevant to drug discovery applications. Without venturing too far into details of system setup, which is beyond the scope of this article, there are some points to be stressed. 

@@

The calculations aim to capture the alchemical change from one ligand to another as accurately as possible. Therefore, it is necessary to consider details of the experimental setup, such as pH. Biological assays are usually run at neutral pH but this is not always the case, and protein and ligand preparation protocols should reflect this. In more detail, the formal charge, or tautomeric state of the small molecules can change within a series of analogs. The change of pKa is a common medicinal chemistry strategy to modify drug properties and would normally require explicit effort to incorporate into alchemical calculations. Further concerns about the experimental setup require understanding what protein construct is used in the bioassay. For instance, does the X-ray that is to be used for the calculations match the construct used for screening (i.e. catalytic vs full length, monomer vs dimer etc) (CITE https://www.nature.com/articles/s41598-018-23039-5)? Also, were certain co-factors or partner proteins required in the bioassay? 

@@

Running binding FE calculations in a drug discovery application will typically require the use of software or tools to facilitate the calculations of many compounds. The commercial implementation from Schr\"{o}dinger, FEP+, uses a graphical user interface and once the user has calculated any missing force field parameters, calculations can be prepared and submitted within minutes. The software builds the network of perturbations, connecting similar compounds based on LOMAP methodology~\cite{liu2013lead}, all inputs for the Desmond simulations are built automatically and the job can be saved or run immediately on available GPUs. The cost of this software is often prohibitive but alternative tools are becoming available to allow relatively straight forward running of alchemical FE calculations with other software~\cite{gapsys2015pmx, loeffler2015fesetup} (CITE https://pubs.acs.org/doi/full/10.1021/acs.jcim.9b00105, cite https://pubs.rsc.org/en/content/articlehtml/2020/sc/c9sc03754c, cite https://pubs.acs.org/doi/abs/10.1021/acs.jcim.5b00368, cite https://jcheminf.biomedcentral.com/articles/10.1186/s13321-019-0348-5). One consideration at this point is to define the perturbations suitable for the application. For instance, if the goal is to accurately assess the relative binding energy of a small number of compounds, possibly with challenging synthesis, the map of perturbations may contain as many connections between compounds as affordable. However, when running calculations on hundreds of compounds a so called ‘star-map’ can be used that just contains one connection per compound: perturbing every compound to a central ligand, typically the crystal structure ligand (CITE https://pubs.acs.org/doi/abs/10.1021/acs.jcim.9b00367). In this way the top-ranking examples can be readily identified and submitted to additional calculations in a second round. 
%
\subsection*{Making predictions, understanding errors}
For prospective drug discovery applications there are several other considerations including understanding likely errors and taking selection bias into account. It is crucial when proposing compounds for synthesis to have some idea of the underlying error or uncertainty in the predictions. A retrospective assessment can give an indication of prospective performance for similar molecules (CITE https://pubs.acs.org/doi/abs/10.1021/acs.jcim.6b00220). Beyond this, several parameters provide useful indicators of performance, results suggest MBAR errors for perturbations greater than 0.3 kcal/mol can be indicative of inaccuracy~\cite{perez-benito2019predicting}. In addition, the hysteresis can be checked by comparing the difference for forward and reverse perturbations of the same structural change, or in the case of large maps of perturbations cycle closure hysteresis~\cite{wang2013modeling} indicates problematic perturbations involved in cycles connecting many compounds. Selection bias, as described above, can be problematic in prospective applications as there is normally an aim to focus on only a narrow range of predicted activity (i.e. the most active). Thus, correlation between prediction and subsequent experiment may lead to disappointment but corrections can be applied based on previous recommendations~\cite{abel2017critical}. 
%

In summary, the successful use of alchemical calculations in drug discovery requires working in the domain of applicability, using a high quality X-ray structure of the target compounds, and testing the approach retrospectively to ensure the system setup is well-behaved. Always assess the confidence in the predictions and communicate this when discussing with experimentalists. Consider performing repeat calculations for at least some of the perturbations in the study. In a drug discovery setting it is inevitable to consider larger virtual libraries, in these cases again evaluate if shorter simulation time per $\lambda$ window can be sufficient for a ‘virtual screening’ mode of application. It is expected that future tools may allow efficient exploration of large virtual libraries. There are many accounts of success of alchemical FE calculations, the methods show good performance towards the goal of binding energy prediction. However, it is important to have realistic expectations. Structure based drug design projects are often capable of improving potency relatively quickly and the range of activity narrows to just two-to-three log units. It may seem hard to impact with substantially different, more potent, stand-out compounds in this scenario, but binding energy prediction can still be extremely useful for ensuring activity is maintained as other properties are optimized. An interesting cost benefit analysis has shown the value of activity prediction, see discussion above and articles such as~\cite{mobley2012perspectiv}. From a drug discovery point of view, FEP is expanding its domain of applicability, there are reports of success using homology models and GPCRs for instance, as well as enabling charge change and scaffold hopping, but these systems are undoubtedly more difficult.  In the meantime, the use cases are expanding to resistance prediction, disease mutation prediction, solubility prediction – an exciting future for alchemical FE calculations. 
%
%
%
%%%%%%%%%%%%
% Step 1   %
%%%%%%%%%%%%
%
\section{Simulation prerequisites}
\label{sec:step1}

The alchemical free energy protocols, that are discussed in Section~\ref{sec:step2}, are defined for a specific type of free energy calculation, i.e. a free energy of binding or a free energy of hydration. Depending on the type of simulations different considerations have to be made for ligands, solvent, and host molecules (in the case of the estimation of free energies of binding).
\subsection*{Free energies of binding}
In principle in the limit of sufficient conformational sampling, the free energy changes estimated from an alchemical free energy calculation should be independent of the system's initial coordinates. However, in practice because simulations of thermodynamic states are of finite duration (typically currently 1-100 ns per state) this is only true for certain classes of alchemical free energy calculations such as relative or absolute free energies of hydration of small and relatively rigid organic molecules. Host-guest or protein-ligand complexes typically exhibit slowly relaxing degrees of freedom that exceed significantly the duration of an alchemical free energy calculation. It is therefore generally important to carefully select input coordinates to obtain satisfactory results. 
The following questions may be relevant before diving into the simulation setup. 
\begin{itemize}
    \item Do I have one or multiple good host structures? (e.g. a good resolution X-ray crystal of the protein target)
    \item Do I have information on one or all of the ligands binding sites (e.g. a X-ray structure)
    \item Should I include buried waters, or other small molecules that can be found in an X-ray structure.
    \item Are my ligands part of a congeneric series? (simple R group substitutions around the same scaffold)
    \end{itemize}

@@

Many relative free energy planning tools have been set up to avoid changing the net charge of the systems considered, including LOMAP~\cite{} and Schr\"{o}dinger's FEP+~\cite{chen2018accurate}. Absolute free energy calculations can also potentially avoid changing the charge of the system by making a charge perturbation of equal and opposite sign elsewhere in the system; for example, as a charged ligand is removed, a charged counterion of opposite sign could also be removed, or one of the same sign could be inserted. This is sometimes referred to as an "alchemical ion" approach for dealing with the needed charge change, and is also employed by the Yank free energy package~\cite{wang2013identifying}.

@@

Before launching alchemical free energy calculations it is wise to consider how convergence and completion will be assessed. Different conditions on when to stop alchemical free energy calculations should be determined, and this may require several iterative checks and therefore modifications to the calculation protocol.
One useful metric to use for termination is the expected or desired uncertainty of a desired free energy estimate, though care must be exercised should the uncertainty estimate prove unreliable.
In particular, if the rate of change in the free energy estimate is significant when this condition is met, the simulation may not be locally converged, and more sampling may be necessary to determine a stable free energy estimate which is no longer changing significantly over time. 
However, this is not the only metric which should be used, as the uncertainty only captures the information about the sampled phase space, not necessarily the entirety of the phase space.  
For example, 
convergence of relative free energy calculations in predictive simulations where the entire phase space is not known in advance, requires sampling the different kinetically stable states. \todo[inline]{cite Mobley and Klimovich JCP 2012} 
This highlights the importance of choosing the correct thermodynamic path to ensure you sample the required thermodynamic states as discussed in Section~\ref{sec:important_path}.

@@

\paragraph{Subsampling data to generate uncorrelated samples}
Once the statistical inefficiency $g$ has been estimated, it is straightforward to subsample the correlated timeseries simulation data to produce effectively uncorrelated data that can be fed to estimators of the free energy.
Suppose the correlated timeseries is $\{a_t\}_{t=1}^T$; we can form a new timeseries of $N_{\mathrm{eff}} \approx T / g$ effectively uncorrelated samples by selecting a subset of indices $\{ \: t = \mathrm{round}((n-1) \, g) \: | \: n \in \mathrm{range}(1,\ldots ,N) \: \}$ where $\mathrm{round(x)}$ denotes rounding to the nearest integer.
